Sleep, relative to wake, increases both veridical and false memory in the DRM paradigm:

### 2023-07-10 manuscript.Rmd
Fixed semantic distance analysis

packagelist <- c('bookdown','knitr','markdown','osfr','plyr','tidyr','dplyr','emmeans','ggplot2','Rmisc','buildmer','lme4','lmerTest','stringr','effects','effectsize','ggdist','car','gridExtra','kableExtra','rstatix','LSAfun','simr')
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

@@

dev.off()
}

@@

if (buildfigures==1){
  pdf('Figures/studiedgraph.pdf',width=15,height=10)

@@

dev.off()

@@

@@

# processdata <- 1

@@

lure_item <- head(fm$word, 20)

@@

@@

semsum <- summary(semantic_ANOVA <- aov(meanDistance~Delay*Test.Time, data=average_closest))
semmeans <- emmeans(semantic_ANOVA, pairwise ~ Test.Time | Delay)

save(file='data/processeddata.RData',list=c('intrusion_poisson','intrusionout','intrusion_CI','lure_model_buildmer','lure_model','lure_pairwise','lureeffects','lure_model_without_covariate_buildmer','null_lure','lureBF','lure_delay_model1','lure_immediate_model1','lure_delay_model1_null','lure_immediate_model1_null','luredelayBF','lureimmediateBF','lurettest','studied_model_buildmer','studied_model','studiedem1','studiedem2','studiedBF','studied_delay_model1','studied_delay_model1_null','studdelayBF','studied_immediate_model1','studied_immediate_model1_null','studimBF','prem','prsum','iota_buildmer000','iota_buildmer2','iota_buildmer2_standard','iota_buildmer3_standard','my_list','closest','average_closest','semsum','semmeans'))

@@

In light of the finding that participants who completed free recall in the evening (vs. morning) produced more intrusions, we asked whether this was driven by these participants having a greater tendency to put down more responses generally. To test this, we calculated the number of total responses by each participant (i.e., studied + lures + intrusions), which is summarised across groups in Figure 5.

@@

In this exploratory analysis, we followed Diekelmann et al. [-@diekelmann2010a] by focusing on the Delay groups\footnote{Note that Diekelmann et al. (2010) did not have Immediate/Control groups.} and by performing a median split on our sample’s adjusted veridical recall (Median = 5), classifying our participants as either high (>5; N = 117) or low ($\leq$ 5; N = 123) performers. We followed the analysis approach of Diekelmann et al. [-@diekelmann2010a], but since the number of critical lures a participant produced is count data, we used a 2 (Group: Sleep vs. Wake) $\times$ 2 (Adjusted veridical recall: High vs. Low Performers) Poisson regression, instead of an ANOVA. This analysis showed no main effect of Group ($\beta$ = `r round(as.numeric(emout5[7]),digits=3)`, z = `r round(as.numeric(emout5[10]),digits=3)`, p = `r p_format(as.numeric(emout5[11]),digits=3,leading.zero=F)`) but a main effect of Adjusted veridical recall ($\beta$ = -0.438, z = -3.766, p < .001), such that high performers (M = 3.17, SD = 2.71) tended to falsely recall more critical lures than low performers (M = 2.09, SD = 1.68) ([@thapar2001a; @toglia1999], but see [@cann2011a; @roediger2001a; @stadler1999a]). Furthermore, importantly, the interaction between Group and Adjusted veridical recall was not significant ($\beta$ = 0.049, z = 0.30, p = .763). Despite this, we followed Diekelmann et al. [-@diekelmann2010a] by comparing low performers in the Sleep and Wake groups. Contrary to their findings, emmeans showed that our low performers in both groups recalled essentially the same number of critical lures (M$_{Sleep}$ = `r round(premmeans[2,3],digits=2)`, SD$_{Sleep}$ = `r round(premmeans[2,4],digits=2)` vs. M$_{Wake}$ = `r round(premmeans[4,3],digits=2)`, SD$_{Wake}$ = `r round(premmeans[4,4],digits=2)`; $\beta$ = 0.133, z = 0.021, p = .982). In other words, this exploratory analysis found no evidence that the effect of sleep on false recall was modulated by veridical memory, casting doubt over the robustness and reliability of Diekelmann et al.’s [-@diekelmann2010a] findings.

@@

Across the Immediate and Delay groups, the number of intrusions and correct recall per list both had a main effect (ps $\leq$ .001) such that they were positively correlated with lure recall. However, the effect of Test Time was not significant in either group (ps > .11). Finally, for the Correct recall per list $\times$ Test Time interaction, it was significant in the Delay (z = `r round(modout2[5,3],digits=3)`, p `r p_format(modout2[5,4],digits=3,accuracy=0.001,leading.zero=F)`) but not the Immediate groups (z = `r round(modout1[5,3],digits=3)`, p = `r p_format(modout1[5,4],digits=3,accuracy=0.001,leading.zero=F)`). To interpret the former, we used the R package, effects [@fox2022a], to visualise it (see Figure \@ref(fig:delayfig)A) and plotted lure recall probability against veridical recall on a participant level (see Figure \@ref(fig:delayfig)B).

@@

As per the DRM literature, responses that were neither the studied list words nor the critical lures were classified as (non-critical) intrusions. For instance, our participants studied nurse, sick, lawyer, medicine, health, hospital, dentist, and physician, with doctor as the critical lure. Responses such as clinic and coconut would both be considered intrusions, but clearly, clinic is semantically more related to the studied list words than coconut. In other words, there is much diversity within the intrusion data. Here, we explored whether our four groups differed in terms of the semantic distance between intrusions and critical lures, as indexed by pre-trained semantic spaces (ukWaC; [@baroni2008a]) derived from word2vec [@gunther2015a]. We reasoned that since lure recall was greater in our Sleep (vs. Wake) participants, the intrusions produced by these participants could be more related to the lures in semantic space (e.g., @mak2023a). To test this, we computed the cosine similarities between each intrusion and each of the 20 critical lures (see Table \@ref(tab:table7) for an illustration). The intrusion-lure pair with the highest cosine similarity (i.e., the nearest neighbour) was used for this analysis.

Since the number of intrusions produced varies greatly between participants, we averaged the lure-intrusion cosine on a participant level and used this as the dependent variable. A 2 (Interval) $\times$ 2 (Test Time) between-participant ANOVA revealed no effects of Interval (z = 2.541, p = .112) or Test Time (z = 0.085, p = .771), and their interaction was also non-significant (z = .886, p = .347). We explored further by comparing the cosine in the Sleep and Wake groups. While this comparison is in the predicted direction (M$_{Sleep}$ = 0.381, SD$_{Sleep}$ = 0.072 vs. M$_{Wake}$ = 0.370, SD$_{Wake}$ = 0.085), it was not statistically significant according to an emmeans pairwise comparison (z = 0.874, p = .383). We interpret these null findings as suggesting that the effect of sleep in the DRM paradigm may be fairly restricted to the lures/studied list words.

@@

**Mixed evidence in the existing literature**.	Our well-powered registered report had a total sample size of 488, with 120 in each of the Sleep and Wake groups. This far surpasses the sample sizes of prior ‘Sleep $\times$ DRM’ studies, where the median N was 27.6 per group. In our data, the effect of sleep on lure recall was only significant when controlling for differences in intrusion rates, and we checked the size of this sleep effect in an exploratory analysis, putting the estimate at Cohen’s d = 0.274.\footnote{We used the eff\_size function in the emmeans package to estimate the effect size of Test Time (sleep vs. wake) in our confirmatory GLM model, with sigma = sigma(lure\_model) and edf = infinite.} This estimate is substantially lower than that from Newbury and Monaghan’s [-@newbury2019a] meta-analysis, which reported to be Cohen’s d = +0.92 (95% CI: 0.54, 1.30). The larger sample size and well-powered nature of our registered report contribute to a more precise estimation of the effect size, highlighting the need for cautious interpretation of prior evidence and the possibility that prior effect sizes may have been inflated due to, for example, publication bias and small sample sizes. Furthermore, given sleep had such a small effect on DRM false memory, it is perhaps not surprising that previous studies had produced mixed evidence. In light of these, we echo the view of a recent article [@cordi2021a] that future sleep research must prioritise robust methodologies (e.g., registered report and pre-registration) and larger sample sizes to enhance the reliability and generalisability of sleep-related memory effects [see also @nemeth2021a].

@@

@@

The first step is to determine a sample size. We have the financial resources to reach up to 160 participants/group (i.e., 640 in total), so we began by fabricating a dataset containing 120 participants/group. We made up 20 false recall observations for each participant. We then split the dataset by Interval, so one dataset for the Delay (Sleep + Wake) groups, another for the Immediate (AM + PM-control) groups, with each containing 4800 observations from 240 participants. In the first dataset, we simulated the false recall data for the Delay groups such that they approximated the data distribution [M$_{Sleep}$ = 45.9% (SD = 20.6%) vs M$_{Wake}$ = 36.3% (SD = 21.2%), p = .005] from a prior study (Payne et al., [-@payne2009a]; Experiment 1). Then, the data were manipulated to fit with our effect size assumption, such that the effect size for sleep is d = +0.54 [M$_{Sleep}$ = 44.7% (SD = 12.6%) vs. M$_{Wake}$ = 38.0% (SD = 12.0%); t(237.4) = 4.20, p = .001]. Afterwards, we simulated the false recall data in the second dataset for the AM and PM-control groups such that they also approximated the data distribution in Payne et al. [-@payne2009a] (Experiment 1; MAM = 42.5% (SD = 19.6%) vs. MPM = 46.3% (SD = 23.5%), p = .57) and that they did not differ significantly from each other [MAM = 42.9% (SD = 13.0%) vs. MPM= 43.5% (SD = 14%); t(233.3) = -0.37, p = .709, d = -0.05].\footnote{Notably, the standard deviations (SDs) from Payne et al. (2009; Experiment 1) are larger than those in our fabricated datasets. This is because Payne et al. showed only 8 wordlists (i.e., maximum lure recall = 8) while ours will show 20 (i.e., maximum lure recall = 20). To explain why this matters, a more concrete example here is useful: In Payne et al, a participant falsely recalling 3 lures would have a false recall rate of 37.5\% while another recalling 4 lures would have a rate of 50\%. So there is a 12.5\% difference between each successive number. In our fabricated data, recalling 3 lures has a false recall rate of 15\% while 4 lures has a rate of 20\%, so there is a 5\% difference. Therefore, understandably, the SDs in our fabricated datasets are necessarily lower than theirs (by roughly a half).}

@@

Based on the fixed-effects estimates, we estimated the power a sample size of 480 (i.e., 120/group) has for detecting an Interval and Test Time interaction. We conducted Monte Carlo simulation using the “simr” package [@green2016a] in R (see Box 2 for R codes). After 500 simulations, it was estimated that a sample of this size gives 90.1% power (95% CI: 87.03, 92.49) to detect a significant interaction. Then, we estimated the power we have for detecting a simple effect of Test Time within the Delay groups (i.e., Sleep vs. Wake). Following the simulation procedures above, it was estimated that 120 participants/group (i.e., 240 in the Delay groups) will give about 98.8% power (95% CI: 97.41, 99.56). In sum, our power calculation showed that having 120 participants/groups will give ample power (>90%) to detect both an Interval x Test Time interaction and a simple effect of Sleep vs. Wake. Finally, we also estimated that we will have at least 80% power as long as we have >99 participants/group. Therefore, in case we fail to reach our target sample size of 120 participants/group before funding expires but manage to recruit >99/group, our proposed experiment will still have satisfactory power.

@@

>>>>

packagelist <- c('bookdown','knitr','markdown','osfr','plyr','tidyr','dplyr','emmeans','ggplot2','Rmisc','buildmer','lme4','lmerTest','stringr','effects','effectsize','ggdist','car','gridExtra','kableExtra','rstatix','LSAfun','simr','effsize')
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

@@

invisible(dev.off())
}

@@

pdf('Figures/studiedgraph.pdf',width=15,height=10)

@@

invisible(dev.off())

@@

```{r maketotalgraph}

total$Delay <- str_replace_all(total$Delay, "control", "Immediate/Control")
total$Delay <- str_replace_all(total$Delay, "delay", "12-hour delay")
total$Delay <- factor(total$Delay, levels = c("Immediate/Control","12-hour delay"))

total_CI <- Rmisc::summarySE(data = total, measurevar = "total", groupvars = c("Delay", "Test.Time"), conf.interval = 0.95)

pdf('Figures/totalgraph.pdf',width=15,height=10)

plothandle <- ggplot(total%>%
         mutate(Test.Time = case_when(Delay=="Immediate/Control"~Test.Time,
                       Test.Time== "AM" ~ "Sleep",
                       TRUE ~ "Wake")),
       aes(Test.Time, total, fill=Test.Time))+
  ggdist::stat_halfeye(adjust = .9, width = .7, .width = 0, justification = -.2, point_colour = NA, alpha=0.5) + 
  stat_summary(fun=mean, geom = "bar", width=0.3)+
  labs(y="Number of responses", x="", title="Total responses", size=30)+
  scale_y_continuous(expand = c(0,0),limits = c(0,150),breaks=seq(0,150,20))+
  theme_light()+
  theme(legend.position="none",legend.title = element_blank(), axis.title.y = element_text(size=16),plot.title = element_text(size=25, face="bold"), axis.text.x = element_text(size=16),axis.text.y = element_text(size=25),legend.text = element_text(size=25), axis.title=element_text(size=25,face="bold"),strip.text = element_text(size = 25), aspect.ratio =1/1)+ scale_fill_manual(values = c("darkorange", "cornflowerblue", "darkorange", "cornflowerblue" ))+coord_cartesian(ylim = c(0,150))+facet_wrap(~Delay,scales = "free_x")+ geom_dotplot(binaxis = "y", stackdir = "down",dotsize = 0.25)+
  geom_errorbar(width=.1, aes(ymin=total-ci, ymax=total+ci), position = position_dodge(width = 0.2), data=total_CI%>%
         mutate(Test.Time = case_when(Delay=="Immediate/Control"~Test.Time,
                       Test.Time== "AM" ~ "Sleep",
                       TRUE ~ "Wake")))

invisible(dev.off())

@@

@@

# lure_item <- head(fm$word, 20)
lure_item <- unique(material$lure)

@@

intrusion_with_distance <- cbind(intrusion_only, closest)

@@

semantic_ANOVA <- aov(meanDistance~Delay*Test.Time, data=average_closest)
semmeans <- emmeans(semantic_ANOVA, pairwise ~ Test.Time | Delay)

@@

In light of the finding that participants who completed free recall in the evening (vs. morning) produced more intrusions, we asked whether this was driven by these participants having a greater tendency to put down more responses generally. To test this, we calculated the number of total responses by each participant (i.e., studied + lures + intrusions), which is summarised across groups in Figure \@ref(fig:totalfig).

```{r totalfig, fig.cap="Number of total responses (i.e., studied + lure + intrusions), summarised across the four groups. Each dot represents an individual participant, and the error bars represent 95\\% confidence intervals.", fig.align="center", include=TRUE, echo=FALSE}

knitr::include_graphics('Figures/totalgraph.pdf')

@@

In this exploratory analysis, we followed Diekelmann et al. [-@diekelmann2010a] by focusing on the Delay groups\footnote{Note that Diekelmann et al. (2010) did not have Immediate/Control groups.} and by performing a median split on our sample’s adjusted veridical recall (Median = 5), classifying our participants as either high (>5; N = 117) or low ($\leq$ 5; N = 123) performers. We followed the analysis approach of Diekelmann et al. [-@diekelmann2010a], but since the number of critical lures a participant produced is count data, we used a 2 (Group: Sleep vs. Wake) $\times$ 2 (Adjusted veridical recall: High vs. Low Performers) Poisson regression, instead of an ANOVA. This analysis showed no main effect of Group ($\beta$ = `r round(as.numeric(emout5[7]),digits=3)`, z = `r round(as.numeric(emout5[10]),digits=3)`, p = `r p_format(as.numeric(emout5[11]),digits=3,leading.zero=F)`) but a main effect of Adjusted veridical recall ($\beta$ = -0.438, z = -3.766, p < .001), such that high performers (M = 3.17, SD = 2.71) tended to falsely recall more critical lures than low performers (M = 2.09, SD = 1.68) ([@thapar2001a; @toglia1999a], but see [@cann2011a; @roediger2001a; @stadler1999a]). Furthermore, importantly, the interaction between Group and Adjusted veridical recall was not significant ($\beta$ = 0.049, z = 0.30, p = .763). Despite this, we followed Diekelmann et al. [-@diekelmann2010a] by comparing low performers in the Sleep and Wake groups. Contrary to their findings, emmeans showed that our low performers in both groups recalled essentially the same number of critical lures (M$_{Sleep}$ = `r round(premmeans[2,3],digits=2)`, SD$_{Sleep}$ = `r round(premmeans[2,4],digits=2)` vs. M$_{Wake}$ = `r round(premmeans[4,3],digits=2)`, SD$_{Wake}$ = `r round(premmeans[4,4],digits=2)`; $\beta$ = 0.133, z = 0.021, p = .982). In other words, this exploratory analysis found no evidence that the effect of sleep on false recall was modulated by veridical memory, casting doubt over the robustness and reliability of Diekelmann et al.’s [-@diekelmann2010a] findings.

@@

Across the Immediate and Delay groups, the number of intrusions and correct recall per list both had a main effect (ps $\leq$ .001) such that they were positively correlated with lure recall. However, the effect of Test Time was not significant in either group (ps > .11). Finally, for the Correct recall per list $\times$ Test Time interaction, it was significant in the Delay (z = `r round(modout2[5,3],digits=3)`, p `r p_format(modout2[5,4],digits=3,accuracy=0.001,leading.zero=F)`) but not the Immediate groups (z = `r round(modout1[5,3],digits=3)`, p = `r p_format(modout1[5,4],digits=3,accuracy=0.001,leading.zero=F)`). To interpret the former, we used the R package, effects [@fox2019a], to visualise it (see Figure \@ref(fig:delayfig)A) and plotted lure recall probability against veridical recall on a participant level (see Figure \@ref(fig:delayfig)B).

@@

As per the DRM literature, responses that were neither the studied list words nor the critical lures were classified as (non-critical) intrusions. For instance, our participants studied nurse, sick, lawyer, medicine, health, hospital, dentist, and physician, with doctor as the critical lure. Responses such as clinic and coconut would both be considered intrusions, but clearly, clinic is semantically more related to the studied list words than coconut. In other words, there is much diversity within the intrusion data. Here, we explored whether our four groups differed in terms of the semantic distance between intrusions and critical lures, as indexed by pre-trained semantic spaces (ukWaC; [@baroni2009a]) derived from word2vec [@guenther2015a]. We reasoned that since lure recall was greater in our Sleep (vs. Wake) participants, the intrusions produced by these participants could be more related to the lures in semantic space (e.g., @mak2023a). To test this, we computed the cosine similarities between each intrusion and each of the 20 critical lures (see Table \@ref(tab:table7) for an illustration). The intrusion-lure pair with the highest cosine similarity (i.e., the nearest neighbour) was used for this analysis.

``` {r extractsemdist}

aovresults <- unlist(summary(semantic_ANOVA))
temp <- capture.output(semmeans$contrasts[2])
emout6 <- unlist(strsplit(gsub("([*]+$)|[<]", "", temp[2]), "\\s+"))

sdmeans <- ddply(average_closest, .(Delay, Test.Time), summarise, average=mean(meanDistance, na.rm=T), stdev=sd(meanDistance, na.rm = T))

Since the number of intrusions produced varies greatly between participants, we averaged the lure-intrusion cosine on a participant level and used this as the dependent variable. A 2 (Interval) $\times$ 2 (Test Time) between-participant ANOVA revealed no effects of Interval (z = `r round(aovresults[13],digits=3)`, p = `r p_format(aovresults[17],digits=3,accuracy=0.001,leading.zero=F)`) or Test Time (z = `r round(aovresults[14],digits=3)`, p = `r p_format(aovresults[18],digits=3,accuracy=0.001,leading.zero=F)`), and their interaction was also non-significant (z = `r round(aovresults[15],digits=3)`, p = `r p_format(aovresults[19],digits=3,accuracy=0.001,leading.zero=F)`). We explored further by comparing the cosine in the Sleep and Wake groups. While this comparison is in the predicted direction (M$_{Sleep}$ = `r round(sdmeans[3,3],digits=3)`, SD$_{Sleep}$ = `r round(sdmeans[3,4],digits=3)` vs. M$_{Wake}$ = `r round(sdmeans[4,3],digits=3)`, SD$_{Wake}$ = `r round(sdmeans[4,4],digits=3)`), it was not statistically significant according to an emmeans pairwise comparison (z = `r emout6[9]`, p = `r p_format(as.numeric(emout6[10]),digits=3,accuracy=0.001,leading.zero=F)`). We interpret these null findings as suggesting that the effect of sleep in the DRM paradigm may be fairly restricted to the lures/studied list words.

@@

**Mixed evidence in the existing literature**.	Our well-powered registered report had a total sample size of 488, with 120 in each of the Sleep and Wake groups. This far surpasses the sample sizes of prior ‘Sleep $\times$ DRM’ studies, where the median N was 27.6 per group. In our data, the effect of sleep on lure recall was only significant when controlling for differences in intrusion rates, and we checked the size of this sleep effect in an exploratory analysis, putting the estimate at Cohen’s d = 0.274.\footnote{We used the eff\_size function in the emmeans package to estimate the effect size of Test Time (sleep vs. wake) in our confirmatory GLM model, with sigma = sigma(lure\_model) and edf = infinite.} This estimate is substantially lower than that from Newbury and Monaghan’s [-@newbury2019a] meta-analysis, which reported to be Cohen’s d = +0.92 (95% CI: 0.54, 1.30). The larger sample size and well-powered nature of our registered report contribute to a more precise estimation of the effect size, highlighting the need for cautious interpretation of prior evidence and the possibility that prior effect sizes may have been inflated due to, for example, publication bias and small sample sizes. Furthermore, given sleep had such a small effect on DRM false memory, it is perhaps not surprising that previous studies had produced mixed evidence. In light of these, we echo the view of a recent article [@cordi2021a] that future sleep research must prioritise robust methodologies (e.g., registered report and pre-registration) and larger sample sizes to enhance the reliability and generalisability of sleep-related memory effects [see also @nemeth2019a].

@@

```{r powersetup, echo=FALSE, include=FALSE}

fabricated <- read.csv('data/Fabricated_FalseRecall_Data.csv')

## Descriptive statistics 
### within the delay groups
#subset (delay groups only)
delay_only <- subset(fabricated, Interval=="Delay")

#mean per participant
delay_only_pp <- ddply(delay_only, .(participant, Test_Time), summarise, lure_recalled=sum(correct), lure_recalled_percent=(sum(correct)/20)*100)

#t test comparing sleep vs. wake
powert1 <- t.test(lure_recalled~Test_Time, data=delay_only_pp) #p < 0.001

#effect size between sleep and wake
powerd1 <- cohen.d(lure_recalled~Test_Time, data=delay_only_pp) #d = 0.54

#mean for sleep and wake
delay_only_group_level <- ddply(delay_only_pp, .(Test_Time), summarise, average_lure_percent=mean(lure_recalled_percent), stdev_lure_percent=sd(lure_recalled_percent))

### Within the immediate groups

#subset (immediate groups only)
immediate_only <- subset(fabricated, Interval=="Immediate")

#mean per participant
immediate_only_pp <- ddply(immediate_only, .(participant, Test_Time), summarise, lure_recalled=sum(correct), lure_recalled_percent=(sum(correct)/20)*100)

#t test comparing AM vs. PM control
powert2 <- t.test(lure_recalled~Test_Time, data=immediate_only_pp) #p = 0.708

#effect size between AM and PM control
powerd2 <- cohen.d(lure_recalled~Test_Time, data=immediate_only_pp) #d = -0.05

#mean and SD for AM and PM control
delay_only_group_level <- ddply(immediate_only_pp, .(Test_Time), summarise, average_lure_percent=mean(lure_recalled_percent), stdev_lure_percent=sd(lure_recalled_percent))

## Compute 2 (Interval) x 2 (Test Time) generalised linear mixed-effect model

#contrast coding
fabricated$Test_Time <- as.factor(fabricated$Test_Time)
contrasts(fabricated$Test_Time) <- contr.sum(2)
fabricated$Interval <- as.factor(fabricated$Interval)
contrasts(fabricated$Interval) <- contr.sum(2)

#build GLMM using buildmer, which automatically finds the maximal model that is capable of converging
fabricated_model <- buildmer(correct~Interval*Test_Time+(1|participant)+(Interval*Test_Time|item), data=fabricated, family="binomial",buildmerControl = buildmerControl(direction='order', args = list(control=glmerControl(optimizer="bobyqa")))) 
#fabricated_model has a by-participant intercept only.

#This is just to compute the same thing again with lmer so that we can run power analysis using simr, which is not compatible with buildmer.
fabricated_model <- glmer(correct~Interval*Test_Time+(1|participant), data=fabricated, family="binomial", control=glmerControl(optimizer="bobyqa"))

#fixed-effect estimates in Table 3
summary(fabricated_model)

fixef(fabricated_model)["Interval1:Test_Time1"] <- 0.076
set.seed(99)
psout <- powerSim(fabricated_model, fixed("Interval1:Test_Time1"), nsim=500, progress=FALSE)

### Simple effect of sleep vs. wake
#compute GLMM within the delay groups
fabricated_simple_model <- glmer(correct~Test_Time+(1|participant), data=delay_only, family="binomial", control=glmerControl(optimizer="bobyqa"))

#fixed effects estimate from the model above
summary(fabricated_simple_model)

#beta for Test Time = 0.139
fixef(fabricated_simple_model)["Test_TimePM"] <- 0.139

# Run simulation
set.seed(99)
psout2 <- powerSim(fabricated_simple_model,fixed("Test_TimePM"),nsim=500)

save(file='data/processeddata.RData',list=c('intrusion_poisson','intrusionout','intrusion_CI','lure_model_buildmer','lure_model','lure_pairwise','lureeffects','lure_model_without_covariate_buildmer','null_lure','lureBF','lure_delay_model1','lure_immediate_model1','lure_delay_model1_null','lure_immediate_model1_null','luredelayBF','lureimmediateBF','lurettest','studied_model_buildmer','studied_model','studiedem1','studiedem2','studiedBF','studied_delay_model1','studied_delay_model1_null','studdelayBF','studied_immediate_model1','studied_immediate_model1_null','studimBF','prem','prsum','iota_buildmer000','iota_buildmer2','iota_buildmer2_standard','iota_buildmer3_standard','my_list','closest','average_closest','semantic_ANOVA','semmeans','my_list','psout','psout2'))

@@

The first step is to determine a sample size. We have the financial resources to reach up to 160 participants/group (i.e., 640 in total), so we began by fabricating a dataset containing 120 participants/group. We made up 20 false recall observations for each participant. We then split the dataset by Interval, so one dataset for the Delay (Sleep + Wake) groups, another for the Immediate (AM + PM-control) groups, with each containing 4800 observations from 240 participants. In the first dataset, we simulated the false recall data for the Delay groups such that they approximated the data distribution [M$_{Sleep}$ = 45.9% (SD = 20.6%) vs M$_{Wake}$ = 36.3% (SD = 21.2%), p = .005] from a prior study (Payne et al., [-@payne2009a]; Experiment 1). Then, the data were manipulated to fit with our effect size assumption, such that the effect size for sleep is d = +`r round(powerd1$estimate,digits=2)` [M$_{Sleep}$ = 44.7% (SD = 12.6%) vs. M$_{Wake}$ = 38.0% (SD = 12.0%); t(`r round(powert1$parameter,digits=1)`) = `r round(powert1$statistic,digits=2)`, p = `r p_format(powert1$p.value,leading.zero=F,digits=3,accuracy=0.001)`]. Afterwards, we simulated the false recall data in the second dataset for the AM and PM-control groups such that they also approximated the data distribution in Payne et al. [-@payne2009a] (Experiment 1; MAM = 42.5% (SD = 19.6%) vs. MPM = 46.3% (SD = 23.5%), p = .57) and that they did not differ significantly from each other [MAM = 42.9% (SD = 13.0%) vs. MPM= 43.5% (SD = 14%); t(`r round(powert2$parameter,digits=1)`) = `r round(powert2$statistic,digits=2)`, p = .709, d = `r round(powerd2$estimate,digits=2)`].\footnote{Notably, the standard deviations (SDs) from Payne et al. (2009; Experiment 1) are larger than those in our fabricated datasets. This is because Payne et al. showed only 8 wordlists (i.e., maximum lure recall = 8) while ours will show 20 (i.e., maximum lure recall = 20). To explain why this matters, a more concrete example here is useful: In Payne et al, a participant falsely recalling 3 lures would have a false recall rate of 37.5\% while another recalling 4 lures would have a rate of 50\%. So there is a 12.5\% difference between each successive number. In our fabricated data, recalling 3 lures has a false recall rate of 15\% while 4 lures has a rate of 20\%, so there is a 5\% difference. Therefore, understandably, the SDs in our fabricated datasets are necessarily lower than theirs (by roughly a half).}

@@

Based on the fixed-effects estimates, we estimated the power a sample size of 480 (i.e., 120/group) has for detecting an Interval and Test Time interaction. We conducted Monte Carlo simulation using the “simr” package [@green2016a] in R (see Box 2 for R codes). After 500 simulations, it was estimated that a sample of this size gives `r round(100*psout$x/psout$n,digits=1)`% power (95% CI: 87.03, 92.49) to detect a significant interaction. Then, we estimated the power we have for detecting a simple effect of Test Time within the Delay groups (i.e., Sleep vs. Wake). Following the simulation procedures above, it was estimated that 120 participants/group (i.e., 240 in the Delay groups) will give about `r round(100*psout$x/psout$n,digits=1)`% power (95% CI: 97.41, 99.56). In sum, our power calculation showed that having 120 participants/groups will give ample power (>90%) to detect both an Interval x Test Time interaction and a simple effect of Sleep vs. Wake. Finally, we also estimated that we will have at least 80% power as long as we have >99 participants/group. Therefore, in case we fail to reach our target sample size of 120 participants/group before funding expires but manage to recruit >99/group, our proposed experiment will still have satisfactory power.

@@

```{r box2b, include=TRUE, eval=TRUE, echo=FALSE}

psout
